---
output:
  html_notebook: default
---

```{r}
library(baseballDBR)
library(Lahman)
library(tidyverse)
library(GGally)
```

### Cleaning and augmenting Postseason

First, players who appeared in multiple postseason rounds in a year have a row for each round. Which isn't what we want. So we have to sum the counts from the multiple rounds.

```{r}
# BattingPost <- Lahman::BattingPost
#BattingPost <- filter(Lahman::BattingPost, playerID=="ripkeca01" | playerID=="anderbr01")

selected_postseason <- select(BattingPost, -c(round, teamID, lgID))
grouped_postseason <- group_by(selected_postseason, yearID, playerID)
summarized_postseason <- summarize_all(grouped_postseason, sum)
nrow(summarized_postseason)
```
There are 9479 player-year rows in postseason after collapsing a player's multiple postseason series into 1 row.

Now let's see how prevalent NA is in the postseason.

```{r drop_all_nas, warning=TRUE}
nrow(summarized_postseason %>% na.omit())
```
We lose 201 rows if we throw out all the rows with na. But dropping all na's could leave us with a biased dataset if certain types of players have NAs when others don't, so instead let's look at when the na's stop:

```{r}
summarise_all(summarized_postseason, funs(mean(is.na(.))))
```

Only 1884 -> 1892 have NAs, which are for CS, HBP, SH, SF, GIDP. 1903 and forward are complete.

```{r}
postseason_complete_rows <- filter(summarized_postseason, yearID > 1902)
nrow(postseason_complete_rows)
```

This is the same number as just dropping all the na's, from the code chunk named drop_all_nas above, which means there are 0 years when some players have na's but others don't.

```{r}
nrow(postseason_complete_rows)
summarized_postseason_no_na <- na.omit(postseason_complete_rows)
nrow(summarized_postseason_no_na)
```

If those two numbers don't match then the earlier 1903 filtering isn't good enough.

```{r}
summarized_postseason_no_na_at_least_1_AB <- filter(summarized_postseason_no_na, AB > 0)
nrow(summarized_postseason_no_na_at_least_1_AB)
summarized_postseason_no_na$P_OPS <- OPS(summarized_postseason_no_na)
summarized_postseason_no_na$P_RC2002 <- RC2002(summarized_postseason_no_na)
summarized_postseason_no_na <- na.omit(summarized_postseason_no_na)
nrow(summarized_postseason_no_na)
```

If those two numbers match then the only reason players have NA after adding OPS and RC is because they had 0 AB.

So we have 6736 postseason rows from 1903 forward. This is on the low side considering our regular season stats aren't complete anywhere close to 1903.

### Cleaning regular season

```{r}
nrow(Batting)
```

How are traded players represented? Willie Mays was traded once midseason.
```{r}
mays <- filter(Lahman::Batting, playerID=="mayswi01", yearID > 1970)
mays
```

The player is only going to play in the postseason for his last team/league (right?), so let's only keep last stint, throw out the rest.

```{r}
group_by(mays, playerID, yearID) %>% filter(stint == max(stint))
```

Ok, that code works for Willie Mays. Now apply it to the whole Batting table and ensure there is only one row per player-year.
```{r}
Batting <- group_by(Lahman::Batting, playerID, yearID)
nrow(Batting)
only_last_stint <- Batting %>% filter(stint == max(stint))
Batting_with_rows <- summarize(Batting, rows = n())
nrow(only_last_stint)
nrow(Batting_with_rows)
```

If last two numbers equal, then we're all good so far. And looks like 102816 - 95250 =~ 7500 players played for multiple teams in one season.


```{r}
last_stint_grouped_by_year <- ungroup(only_last_stint) %>% group_by(yearID)
summarise_all(last_stint_grouped_by_year, funs(sum(is.na(.))))
```

So it's IBB that we don't have data for most recently. Which is unfortunate. SF and CS probably don't matter that much, but I suspect IBB is a fairly unique and informative signal.

```{r}
nrow(inner_join(only_last_stint %>% filter(yearID > 1954), summarized_postseason_no_na, by=c("playerID", "yearID")))
nrow(inner_join(only_last_stint %>% filter(yearID > 1938), summarized_postseason_no_na, by=c("playerID", "yearID")))
nrow(inner_join(only_last_stint %>% filter(yearID > 1912), summarized_postseason_no_na, by=c("playerID", "yearID")))
```
We still have 5109 rows if we cut off at 1954. If we drop GIDP before 1939 we have 5666, if we drop GIDP before 1913, we have 6480.

```{r}
aug_bat <- inner_join(only_last_stint %>% filter(yearID > 1954), summarized_postseason_no_na, by=c("playerID", "yearID"))
aug_bat <- ungroup(ungroup(aug_bat))
```


```{r}
aug_bat$lgID = ifelse(aug_bat$lgID=="AL",1,0)
aug_bat$playerID = as.numeric(as.factor(aug_bat$playerID))
aug_bat$teamID = as.numeric(as.factor(aug_bat$teamID))
cols_to_drop <- grep("\\.y$", names(aug_bat), value = TRUE)
fin_bat <- select(aug_bat, -ends_with(".y"))
write.csv(fin_bat, file = "training.csv", row.names=FALSE)
```
